Bird beak's elaboration and innovation

### 2023-05-15 Analysis/14-Elab_innov_through_time.Rmd
Update 14-Elab_innov_through_time.Rmd

## Table that for easy plotting
super_orders_elaboration <- cbind(super_orders_elaboration, super_age_info, cex = 2)
super_orders_innovation <- cbind(super_orders_innovation, super_age_info, cex = 2)
orders_elaboration <- cbind(orders_elaboration, order_age_info, cex = 1)
orders_innovation <- cbind(orders_innovation, order_age_info, cex = 1)
elaboration_plot_data <- rbind(super_orders_elaboration, orders_elaboration)
innovation_plot_data <- rbind(super_orders_innovation, orders_innovation)

## Empty plot
plot.age.data <- function(data, ylab, xlab) {
  plot(NULL, xlab = xlab, ylab = ylab,
       xlim = rev(range(pretty(range(data$age)))),
       ylim = range(pretty(range(data[,c(4, 8)]))))
  ## Adding the lines for each clade
  for(i in 1:nrow(data)) {
    lines(x = rep(data[i, "age"][[1]], 2), y = data[i, c("2.5%", "97.5%")], lty = 2, col = data[i, "col"][[1]])
    lines(x = rep(data[i, "age"][[1]], 2), y = data[i, c("25%", "75%")], lty = 1, lwd = 2, col = data[i, "col"][[1]])
    points(x = data[i, "age"][[1]], y = data[i, "obs.median"], cex = data[i, "cex"], col = data[i, "col"][[1]], pch = 19)
  }
}
pdf(height = 14, width = 7, file = "../Manuscript/Figures/Elaboration_and_innovation_through_time.pdf")
par(mfrow = c(2,1))
plot.age.data(elaboration_plot_data, ylab = "Elaboration", xlab = "")
legend("topleft", pch = 19, col = "grey", legend = c("super orders", "orders"), pt.cex = c(2, 1))
plot.age.data(innovation_plot_data, ylab = "Innovation", xlab = "Time (Mya)")
dev.off()
```

>>>>

## Table that for easy plotting
super_orders_elaboration <- cbind(super_orders_elaboration, super_age_info, pch = 19)
super_orders_innovation <- cbind(super_orders_innovation, super_age_info, pch = 19)
orders_elaboration <- cbind(orders_elaboration, order_age_info, pch = 21)
orders_innovation <- cbind(orders_innovation, order_age_info, pch = 21)
elaboration_plot_data <- rbind(super_orders_elaboration, orders_elaboration)
innovation_plot_data <- rbind(super_orders_innovation, orders_innovation)

## Empty plot
plot.age.data <- function(data, ylab, xlab, main = "") {
  plot(NULL, xlab = xlab, ylab = ylab, main = main,
       xlim = rev(range(pretty(range(data$age)))),
       ylim = range(pretty(range(data[c("2.5%", "25%", "75%", "97.5%")]))))
  ## Adding the lines for each clade
  for(i in 1:nrow(data)) {
    lines(x = rep(data[i, "age"][[1]], 2), y = data[i, c("2.5%", "97.5%")], lty = 2, col = data[i, "col"][[1]])
    lines(x = rep(data[i, "age"][[1]], 2), y = data[i, c("25%", "75%")], lty = 1, lwd = 2, col = data[i, "col"][[1]])
    points(x = data[i, "age"][[1]], y = data[i, "obs.median"], pch = data[i, "pch"], col = data[i, "col"][[1]])
  }
}

pdf(height = 14, width = 7, file = "../Manuscript/Figures/Elaboration_and_innovation_clade_through_time.pdf")
par(mfrow = c(2,1))
plot.age.data(elaboration_plot_data, ylab = "Elaboration", xlab = "", main = "Clade's elaboration/innovation\nthrough time")
legend("topleft", pch = 21, col = unlist(orders_elaboration$col), legend = rownames(orders_elaboration), cex = 0.7)
legend("top", pch = 19, col = unlist(super_orders_elaboration$col), legend = rownames(super_orders_elaboration), cex = 0.7)
plot.age.data(innovation_plot_data, ylab = "Innovation", xlab = "Time (Mya)")
dev.off()
```

# Checking elaboration and innovation values through time (species)

```{r, eval = TRUE, echo = TRUE}
## Get all the statistics
super_orders_elaboration <- summary(results$super_results_within$position)[-9,]  #-9 is for removing the "phylogeny" element
super_orders_innovation <- summary(results$super_results_within$distance)[-9,]
orders_elaboration <- summary(results$order_results_within$position)[-28,] #-28 is for removing the "phylogeny" element
orders_innovation <- summary(results$order_results_within$distance)[-28,]

super_age_info <- t(sapply(unlist(strsplit(super_orders_elaboration$subsets, split = ":phylogeny"))[-9], get.clade.age, info = clade_info$super_orders, tree = ploting_data$consensus_tree)) #-9 is for removing the "phylogeny" element
order_age_info <- t(sapply(unlist(strsplit(orders_elaboration$subsets, split = ":phylogeny"))[-28], get.clade.age, info = clade_info$orders, tree = ploting_data$consensus_tree)) #-28 is for removing the "phylogeny" element

## Table that for easy plotting
super_orders_elaboration <- cbind(super_orders_elaboration, super_age_info, pch = 19)
super_orders_innovation <- cbind(super_orders_innovation, super_age_info, pch = 19)
orders_elaboration <- cbind(orders_elaboration, order_age_info, pch = 21)
orders_innovation <- cbind(orders_innovation, order_age_info, pch = 21)
elaboration_plot_data <- rbind(super_orders_elaboration, orders_elaboration)
innovation_plot_data <- rbind(super_orders_innovation, orders_innovation)

pdf(height = 14, width = 7, file = "../Manuscript/Figures/Elaboration_and_innovation_species_through_time.pdf")
par(mfrow = c(2,1))
plot.age.data(elaboration_plot_data, ylab = "Elaboration", xlab = "", main = "Species's elaboration/innovation\nthrough time (per clade)")
legend("topright", pch = 21, col = unlist(orders_elaboration$col), legend = unlist(orders_elaboration$subsets), cex = 0.7)
legend("top", pch = 19, col = unlist(super_orders_elaboration$col), legend = unlist(super_orders_elaboration$subsets), cex = 0.7)
plot.age.data(innovation_plot_data, ylab = "Innovation", xlab = "Time (Mya)")
dev.off()
```
